High Resolution Identification of Protein-DNA Binding Events and Quality Control for ChIP-exo data

### 2015-11-24 rmarkdown/slides.Rnw
finished slides

\documentclass[table,xcolor=dvipsnames]{beamer}

@@

\usetheme{Szeged}
\usecolortheme[named=RoyalBlue]{structure}

@@

\begin{frame}[plain]
  \frametitle{Comparison of ChIP-exo and ChIP-seq - SCC}  
\begin{figure}[H]
  \centering  
\includegraphics[width = .7\textwidth]{../figs/for_paper/scc_ctcf.pdf}
\caption{SCC for CTCF factor in HeLa cell line for ChIP-exo and
  SET-ChIP-Seq }
  \label{fig:scc}
\end{figure}

@@

\begin{frame}
  \frametitle{Comparison of ChIP-exo and ChIP-Seq}

@@

\begin{itemize}
\item
\item 
\end{itemize}

@@

\begin{frame}[t]
\frametitle{Software}  
\begin{itemize}
\item \textbf{dPeak}: We updated the initialization strategy. The
  latest version is currently available from
  \url{http://dongjunchung.github.io/dpeak/}.
\item \textbf{ChIPexoQual}: This package contains the QC pipeline for
  ChIP-exo. The last version is available in
  \url{https://github.com/welch16/ChIPexoQual}.
\item \textbf{Segvis}: The goal of this package is to visualize
  genomic regions by using aligned reads. The latest version is
  available in \url{https://github.com/keleslab/Segvis}.
\item \textbf{ChIPUtils}: This package attempts to gather the most
  commonly used ChIP-Seq QC.The latest available version is in
  \url{https://github.com/welch16/ChIPUtils}.
\end{itemize}

>>>>

\documentclass[compress,table,xcolor=dvipsnames]{beamer}

@@

\usetheme{Szeged}
\usecolortheme[named=RoyalBlue]{structure}
\setbeamertemplate{headline}{}
\setbeamertemplate{navigation symbols}{}

@@

\begin{frame}
  \frametitle{Comparison of ChIP-exo and ChIP-seq I}  
\begin{figure}[H]
  \centering  
\includegraphics[width = .7\textwidth]{../figs/for_paper/scc_ctcf.pdf}
\caption{SCC for CTCF factor in HeLa cell line for ChIP-exo and
  SET-ChIP-Seq }
  \label{fig:scc}
\end{figure}

@@

\begin{frame}
  \frametitle{Comparison of ChIP-exo and ChIP-Seq II}

@@

{\small
\begin{itemize}
\item A shows that high dnesity regions are similar between ChIP-Seq
  and ChIP-exo but background regions are not.
\item The peak-pair assumption doesn't hold in ChIP-exo data, some
  regions show strand-imbalance
\end{itemize}
}

@@

\section{The ChIP-exo QC pipeline}

\begin{frame}[plain]
  \frametitle{The ChIP-exo QC pipeline}

\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2.pdf}
\end{figure}

{\small
  \begin{itemize}
  \item $\mbox{ARC} = \frac{\text{Nr. of reads in the
        region}}{\text{Width of the region}}$
  \item $\mbox{URCR} = \frac{\text{Nr. of reads mapped to only one
        position in the region}}{\text{Nr. of reads in the region}}$
  \item $\mbox{local-NSC}$
  \item $\mbox{FSR} = \frac{\text{Nr. of fwd. strand reads in region}}{\text{Nr. of reads in region}}$
  \end{itemize}  
}

\begin{frame}[t,plain]
  \frametitle{Library complexity and enrichment}
\only<1>{
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
\end{figure}
}

\only<2>{
{\small
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth]{../figs/for_paper/local_SCC_example.pdf}
\end{figure}
}}
\only<2>{
\begin{align}
 y(\delta) = f(x_\delta) + \epsilon_\delta \quad 
\mbox{local-NSC} = \frac{max_{x_\delta} \hat{f}(x_\delta)}{\hat{\sigma}_f}
\nonumber
\end{align}
}

\only<3>{
\begin{figure}[H]
\centering
  \includegraphics[width = \textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
\end{figure}
}

\only<3>{
{\small
where:
\begin{itemize}
\item high - regions with nr. of unique positions $ > 100$
\item med - regions with nr. of unique positions in $(50,100)$
\item low - regions with nr. of unique positions in $(20,100)$
\end{itemize}
}}

\begin{frame}[t]
  \frametitle{Strand imbalance}

\begin{figure}[H]
  \centering  
  \includegraphics[width = \textwidth,page = 3]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
  \label{fig:strand}
\end{figure}

\section{Comparison with ChIP-Seq using dPeak}
\begin{frame}[t]
  \frametitle{dPeak model for SET case}

\only<1>{
  We consider a region with $n$ reads and $m$ positions, for the
  $i$-th read:
  \begin{itemize}
  \item $Z_i \sim \mbox{Multi}(\pi_0,\pi_1,\cdots,\pi_{g^*})$
  \item $D_i \sim \mbox{Ber}(p_D)$
    \begin{itemize}
    \item The read is in the forward strand ($D_i = 1$):
      \begin{itemize}
      \item The reads belongs to the background: $R_i | Z_i = 0, D_i =
        1 \sim \mbox{Unif}(1 - \beta + 1, m)$
      \item The read belong to the $g$-th binding event: $R_i | Z_i =
        g, D_i = 1 \sim \mbox{N}(\mu_g - \delta , \sigma^2)$
      \end{itemize}
    \item The read is in the backward strand ($D_i = 0$):
      \begin{itemize}
      \item The reads belongs to the background: $R_i | Z_i = 0, D_i =
        0 \sim \mbox{Unif}(1 , m + \beta - 1)$
      \item The read belong to the $g$-th binding event: $R_i | Z_i =
        g, D_i = 0 \sim \mbox{N}(\mu_g + \delta , \sigma^2)$
      \end{itemize}
    \end{itemize}
  \end{itemize}
}

\only<2>{
      \begin{figure}[H]
        \centering
        \includegraphics[width =
        .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/poster/dpeak.png}
      \end{figure}
}
\end{frame}

\begin{frame}
\frametitle{Comparison with ChIP-Seq using dPeak}
\only<1>{
  \begin{figure}[H]
  \centering
  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
\end{figure}
}
\only<1>{
{\small
  \begin{itemize}
  \item Sensitivity is the defined as the proportion of identified
    peaks (regulonDoB \cite{regulonDB} is used as gold-stantard)
  \item Resolution is defined as the min. absolute distance of a
    regulonDB annotation to an est. binding location.
  \end{itemize}
}}

\only<2>{
\begin{figure}[H]
\centering
   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf} 
 \end{figure}
}
\only<2>{
{\small
\begin{itemize}
\item $\delta$ measures the average distance of reads to their
  respective binding sites
\item $\sigma$ measures the dispersion of reads around their
  respective binding sites
\end{itemize}
}}
\end{frame}

\begin{frame}[t]
  \frametitle{ChIP-Seq comparison at fixed depth}

  We sampled $n$ fragment reads of each dataset ($2n$ for PET
  ChIP-Seq), and applied the MOSAiCS / dPeak pipeline:

\only<1>{
\begin{figure}[H]
  \centering
  \includegraphics[width = .4\textwidth,page = 1]{../figs/for_paper/saturation_analysis_old_frames.pdf}
  \includegraphics[width = .4\textwidth,page = 2]{../figs/for_paper/saturation_analysis_old_frames.pdf}
\end{figure}
}
\only<1>{
{\small
  \begin{itemize}
  \item ChIP-exo and PET ChIP-Seq are comparable and outperform SET
    ChIP-Seq
  \end{itemize}
}}

\only<2>{
\begin{figure}[H]
\centering
\includegraphics[width = .4\textwidth,page = 3]{../figs/for_paper/saturation_analysis_old_frames.pdf}
  \includegraphics[width = .4\textwidth,page = 4]{../figs/for_paper/saturation_analysis_old_frames.pdf}
\end{figure}
}

\section{Conclusions and future work}

\begin{frame}
  \frametitle{Conclusions}

  \begin{itemize}
  \item Our pipeline is capable of assessing the balance between
    sample enrichment and library complexity.
  \item We shown that the ``peak-pair'' assumption doesn't hold well
    in practice, and implemented a visualization capable of detecting
    strand imbalance.
  \item We updated dPeak, which makes a striking balance in
    sensitivity, specificity and spatial resolution.
  \item ChIP-exo and PET ChIP-Seq are comparable in resolution and
    sensitivity, and both outperform SET ChIP-Seq.
  \item We showed that with a fixed number of reads, ChIP-exo
    outperforms PET and SET ChIP-Seq.
  \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Future work}

\begin{itemize}
\item In the paper, we showed that there is a relationship between
  ChIP-exo tag counts and both mappability and GC content scores. We
  want to add a QC measure to the pipeline based on them.
\item We want to assess if ChIP-Nexus library complexity is actually higher than ChIP-exo's by using the $\mbox{local-NSC}$.
\item We have been studying E. Coli's transcription initiation
  complexes with PET ChIP-Seq, being able to label regions as open or
  closed complexes. We want to improve this analysis by using ChIP-exo
  data, and hopefully detecting intermediate step between this two
  states.
\item Find a optimal strategy for labelling enhancer out of a
  predetermined list of regions in the genome by the use of active
  learning techniques.
\end{itemize}

\begin{frame}[t]
\frametitle{Software}  
\begin{itemize}
\item {\color{RoyalBlue}\textbf{dPeak}}: We updated the initialization
  strategy. The latest version is currently available from
  \url{http://dongjunchung.github.io/dpeak/}.
\item {\color{RoyalBlue}\textbf{ChIPexoQual}}: This package contains
  the QC pipeline for ChIP-exo. The last version is available in
  \url{https://github.com/welch16/ChIPexoQual}.
\item {\color{RoyalBlue}\textbf{Segvis}}: The goal of this package is
  to visualize genomic regions by using aligned reads. The latest
  version is available in \url{https://github.com/keleslab/Segvis}.
\item {\color{RoyalBlue}\textbf{ChIPUtils}}: This package attempts to
  gather the most commonly used ChIP-Seq QC.The latest available
  version is in \url{https://github.com/welch16/ChIPUtils}.
\end{itemize}

\begin{frame}[plain]

{\Huge
{\color{RoyalBlue}
\begin{center}
  \textbf{Thank you very much!}
\end{center}
}
}
